\textsc{RILS}-\textsc{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-27 paper_resources/doc/paper.tex
Section 4.4.

In order to check the statistical significance of the results w.r.t.\ obtained exact models and their differences between the competitor approaches, we employed the statistical methodology comparing the results of \textsc{RILS}--\textsc{ROLS} to 14 other approaches over all problem instances from the both ground-truth benchmark sets, \textsc{Feynman} and \textsc{Strogatz}. Note that an indicator random variable is involved per each run of the algorithms. Thus, the score that comes into the statistical observation is an indicator (0 or 1) indicating weather or not an algorithm was able to return the exact model per each run. More precisely, we involved 1300 results into our statistical evaluation for each of the algorithms. Note that  results of some of the 14 competitors are not known from \url{https://cavalab.org/srbench/results/} due to some unknown reasons (most likely algorithm is terminated running out of memory or due to the chosen  value of an internal parameter). For these runs, we simply count a 0 for the score.

Initially, Friedman’s test was separately executed for all competitor approaches.      In those cases in which the null hypothesis $H_0$ was rejected ($H_0$ states that there are no statistical differences between the obtained results of all competitors) pairwise comparisons are further performed by using the Nemenyi post-hoc test~\cite{pohlert2014pairwise}. The outcome is represented by means of critical difference (CD) plots. In each CD plot, the (15) competitor approaches are placed on the horizontal axis according to their average ranking. Thereafter, the CD score is computed for a significance level of 0.05. If the difference is small enough, meaning that no statistical difference is detected, a horizontal bar linking statistically equal approaches is drawn.   Three (one for each level of noise) CD plots are shown in Figure~\ref{fig:cd-plots-exact-pcts} per each level of noise. The following conclusions may be drawn from there.

\begin{itemize}
	\item  Concerning the instance problems where no noise is included, \textsc{RILS}-\textsc{ROLS} method returns the best average ranking in terms of the number of obtained exact models.  The second best average ranking is achieved by \textsc{AI-Feynman} algorithm. There is  statistically significant difference between these two approaches w.r.t.\ number of found exact models. The third best approach is \textsc{Apf-Fe},  which is far behind \textsc{RILS}-\textsc{ROLS} w.r.t.\ delivered average ranking.  
	\item    Concerning the instance problems with the included  level of noise of 0.001, \textsc{RILS}-\textsc{ROLS} method again obtains the best average ranking in terms of the found exact models.   The second best ranking is obtained by \textsc{AI-Feynman} algorithm. There is again a statistically significant difference between the results obtained by these two approaches. The third best approach is \textsc{Afp-Fe} which performs statistically worse than \textsc{AI-Feynman} approach.  
	\item  Concerning the instance problems with the included noise level of  0.01, again, \textsc{RILS-ROLS} method produces the best average ranking according to positively evaluated runs, i.e. the runs with obtained exact models; the second best approach w.r.t.\ average ranking is \textsc{Afp-Fe}, followed by \textsc{Dsr}, and \textsc{Afp}. These latter three approaches perform statistically equal, but significantly worse than \text{RILS}-\textsc{ROLS} method. 	The fourth best approach according to delivered average ranking is \textsc{AI-Feynman} which is on pair with \textsc{gplearn}. Note that the average ranking of \textsc{AI-Feynman} method declines by increasing  the level of noise in the input data. Thus, despite being efficient on no-noisy data, it seems that \textsc{AI-Feynman} is very sensitive to the level of noise increase, which is not that dramatically conspicuous with our \textsc{RILS}-\textsc{ROLS} method.

>>>>

In order to check the statistical significance of the results w.r.t.\ obtained exact models and their differences between the competitor approaches, we employed the statistical methodology comparing the results of \textsc{RILS}--\textsc{ROLS} to 14 other approaches over all algorithm runs of the problem instances from the both ground-truth benchmark sets, \textsc{Feynman} and \textsc{Strogatz}. Note that an indicator random variable is involved per each run of the algorithms. Thus, the score that comes into the statistical observation is an indicator (0 or 1) indicating weather or not an algorithm was able to return the exact model at that run. More precisely, we involved 1300 results into our statistical evaluation for each algorithm. Note that  results of some of the 13 competitors are not known from SRBench \url{https://cavalab.org/srbench/results/} due to   unknown reasons (most likely algorithm is terminated running out of memory or due to   chosen algorithm configuration settings). For these runs, we simply count a 0 for the score.

Initially, Friedman’s test was separately executed for all competitor approaches.      In those cases in which the null hypothesis $H_0$ was rejected ($H_0$ states that there are no statistical differences between the obtained results of all competitors) pairwise comparisons are further performed by using the Nemenyi post-hoc test~\cite{pohlert2014pairwise}. The outcome is represented by means of critical difference (CD) plots. In each CD plot, the (14) competitor approaches are placed on the horizontal axis according to their average ranking. Thereafter, the CD score is computed for a significance level of 0.05. If the difference is small enough, meaning that no statistical difference is detected, a horizontal bar linking statistically equal approaches is drawn.   Three (one for each level of noise) CD plots are shown in Figure~\ref{fig:cd-plots-exact-pcts} per each level of noise. The following conclusions may be drawn from there.

\begin{itemize}
	\item  Concerning the instance problems where no noise is included, \textsc{RILS}-\textsc{ROLS} method returns the best average ranking in terms of the number of obtained exact models over all runs.  The second best average ranking is achieved by \textsc{AI-Feynman} algorithm. There is  statistically significant difference between these two approaches w.r.t.\ number of found exact solutions. The third best approach is \textsc{Apf-Fe},  which is far behind \textsc{RILS}-\textsc{ROLS} w.r.t.\ delivered average ranking.  
	\item    Concerning the instance problems with the included  level of noise of 0.001, \textsc{RILS}-\textsc{ROLS} method again obtains the best average ranking in terms of the found exact solutions.   The second best ranking is obtained by \textsc{AI-Feynman} algorithm. There is again a statistically significant difference between the results obtained by these two approaches. The third best approach is \textsc{Afp-Fe} which performs statistically worse than \textsc{AI-Feynman}.  
	\item  Concerning the instance problems with the included noise level of  0.01, again, \textsc{RILS-ROLS} method produces the best average ranking according to positively evaluated runs, i.e. the runs with obtained exact models; the second best approach w.r.t.\ average ranking is \textsc{Afp-Fe}, followed by \textsc{Dsr}, and \textsc{Afp}. The  latter three approaches perform statistically equal, but significantly worse than \text{RILS}-\textsc{ROLS} method. 	The fourth best approach according to delivered average ranking is \textsc{AI-Feynman} which is on pair with \textsc{gplearn}. Note that the average ranking of \textsc{AI-Feynman} method declines by increasing  the level of noise in the input data. Thus, despite being efficient on the no-noisy data, it seems that \textsc{AI-Feynman} is very sensitive w.r.t.  increase in the level of noise, which is not that dramatically conspicuous with our \textsc{RILS}-\textsc{ROLS} method.
